Penguins Search Optimisation Algorithm for Association Rules Mining

### 2015-02-05 Swarm paper/elsarticle/elsarticle-template-num.tex
some changes are made on Pe-ARM paper

\title{Penguins Search Optimization Algorithm For Association Rules Mining}

@@

\address[b]{Department of Computer Science, University of Hull, Hull HU6 7RX, UK}

\address[a]{ Badji Mokhtar-Annaba University, P.O. Box 12, 23000 Annaba, Algeria}

@@

Association Rules Mining (ARM) is one of the most popular and well-known approaches for the decision-making process. All existing ARM algorithms generate a very large number of association rules with high overlapping. To deal with this issue, we propose a new ARM approach based on penguins search optimization algorithm (Pe-ARM for short). Moreover, an efficient measure is incorporated into the main process to evaluate the amount of overlapping among the generated rules. Indeed, before the evaluation of the rules, each penguin calculates the amount of overlapping between its rule and all accepted rules. Then, the given rule is rejected without fitness evaluation if at least the amount of overlapping for this rule and one rule from the set of accepted rule is greater than the maximum accepted overlap. The proposed approach also ensures a good diversification over the whole solutions space thanks to the multi-groups behavior of penguins. Furthermore, the search process is based on the oxygen reserve of each penguin, which permit to accelerate the search of a good solutions in reasonable time and to save energy in the disagreeable regions. To demonstrate the effectiveness of the suggested approach, several experiments have been carried out on different data-sets and specifically on the biological ones. The results reveal that the proposed approach outperforms the well known ARM algorithms in both execution time and solution quality. 
\end{abstract}

\begin{keyword}
Association Rule Mining; Penguin Search Optimization Algorithm; Overlap measure; Biological Data-set; ARM.
%% keywords here, in the form: keyword \sep keyword

@@

 Association Rules Mining (ARM) is one of the most challenging and important tasks in data mining~\cite{1}. 

@@

the problem solving process. In the literature, ARM algorithms can be divided into two main categories: exact approaches and 
metaheuristics-based approaches. First the exact approaches aim at extracting all possible association rules, such as Apriori~\cite{5}, FPGrowth~\cite{6}, DIC~\cite{7}, DHP~\cite{8} and Eclat~\cite{9}. However, these algorithms are high time and space consuming when dealing with very large databases with many items and transactions. 
To reduce the computational time of such algorithms,second category is metaheuristics-based approaches have been proposed such as: ARMGA~\cite{10} and G3PARM~\cite{11} for evolutionary algorithms, PSO-ARM~\cite{12}, $ACO_R$~\cite{13} and BSO-ARM~\cite{14} 
for swarm intelligence. These algorithms has the potential to give only one part of rules and not all relevant rules. Therefore, two challenges can be derived, the computation time of exact algorithms by exploring all rules space in one hand and the quality of extracted rules using metaheuristics-based approaches by taking into account the similar rules in the other hand. \\
The penguins search optimization algorithm \cite{15} is a recent nature inspired approach based on the collaborative hunting strategies of penguins. The penguins synchronize their dives to reduce the expenditure of energy on the hunting process, this assumption used to help the algorithm to converge quickly to the optimal solution. The diversification strategy of penguins search algorithm allow penguins to explore in an efficient way the whole space solution. The search process of penguins is based on the reserve of oxygen which allow penguins to decide to go fast or no while explorer the search space in one hand and to decide to search or not in a given region on the other hand. This has motivate us to use the PeSOA for the association rule mining problem to explore in efficient way the space solution. We incorporate a new measure to the main process of penguins search to compute the amount of overlapping between rules to help penguins search process to generate only the non redundant rule with low amount of overlapping among them. \\
In this paper a new algorithm called Pe-Arm (Pe-ARM: for association rules mining) is proposed  to find a set of consistent rules having a high quality with low amount of overlapping. These set of rules must cover the maximum number of transaction in the database.\\
The remainder of the paper is organised as follows. In the next section, relevant works on association rule mining with diverse applications are presented. 
Then, the penguins search optimization algorithm is introduced in Section 3, 
followed by the proposed Pe-ARM approach  (penguins search optimization algorithm for association rules mining) in section 4. 
The experimental results with both standard and biological data sets is reported in Section 5. 

@@

Particle swarm optimization algorithm has been applied for association rule mining \cite{12}, 

@@

Other swarm intelligence algorithms have also been applied for ARM problem. First, an Ant colony optimization \cite{20} worked in health application 

@@

a gravitational emulation local search algorithm, which is a nature inspired optimization algorithm. Each rule is modeled as a mass, all the masses attract each others using the law of motion. Each iteration is based on the 

@@

Bees Swarm Optimization (BSO) and Tabu Search (TS). BSO will explore the search space so that it can cover most of its neighbors. 

@@

\section{Penguins Search Optimization Algorithm }

Penguin search optimization algorithm (PeSOA) is a new swarm based metaheuristic algorithm which was proposed in \cite{15}. The dietary behavior of penguins may be explained by the economic reasoning: it comes to a profitable food search activity when the gain of energy is greater than the expenditure required to obtain this gain. Penguins, behaving along the line of foraging predators, must extract information about the time and cost to get food and the energy content of prey in order to choose the course for making their next dive.
 PeSOA is inspired by the penguin’s hunting behavior and it generally works as follows. The population of penguins locates initial positions, this population is divided on a set of groups, and each group is assigned to a region on the whole solution space. Each penguin then dives and swims under the water for hunting fish while consuming its oxygen reserve. Different forms of the communication between penguins are occasionally performed and the quantities of eaten fish increase. The process is repeated until the specified amount of fish is obtained or the maximum number of iterations is reached. The authors of \cite{15} have shown the PeSOA outruns genetic algorithms and particle swarm optimization in obtaining better values for benchmark optimization functions. 

@@

The optimization of ARM aims at maximizing the

@@

rules. Optimization algorithm gives only a set of the pertinent

@@

 After each iteration, the penguins communicate with each other the best rule founded  $GLbest$ to converge for the best group, update the oxygen reserve and the number of neighbours  for each penguin. The computation of the health of each group is made to  define the probabilities of improvement in each group, and finally we redistribute the penguins on the new population according the health of the penguins of each group (probabilities). 
\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=15cm]{2.jpg}
\end{center}
\caption{Penguins Search Optimization Algorithm For Arm}
\label{algo}
\end{figure*}
\section{EXPERIMENTATION AND RESULTS}
We used several evaluation criteria in the experimentation process in order to evaluate the proposed algorithm.
Firstly, the statistical measure is computed which is represented as the average of the confidence and the support of the generated rules. 
Secondly, the run time of each approach is determined. \\
The last measure is the coverage formula which represents the similarities 
between the generated rules according to the number of common transactions. Indeed, 
the rules are similar when they verify together many transactions and dissimilar where
they do not verify any transaction \cite{27}. The coverage formula is given as follows:\\
Let $T_{r_i}$ be the set of transactions verified by $r_{i}$, and $n$ is the number of generated rules.\\
\begin{displaymath}
Coverage =   \frac{\sum_{i=1}^{n}{\sum_{j=1}^{n} {\eta{(r_{i},r_{j})}}}}{n~(n-1)} \\
\end{displaymath}
\begin{displaymath}
\eta{(r_{i},r_{j})} =
\left\lbrace
\begin{array}{ccc}
0          &  i = j & \\
\left|T_{r_{i}} \cup T_{r_{j}}\right| - \left|T_{r_{i}} \cap T_{r_{j}}\right|  & \mbox{Otherwise}  
\end{array}\right.
\end{displaymath}

@@

%% \begin{theorem}[ ] ...  \end{theorem}
%% \begin{definition}[ ] ...  \end{definition}
%% \begin{axiom}[ ] ...  \end{axiom}
%% \begin{conjecture}[ ] ...  \end{conjecture}
%% \begin{proposition}[ ] ...  \end{proposition}
%% \begin{corollary}[ ] ...  \end{corollary}
%% \begin{lemma}[ ] ...  \end{lemma}
%% \begin{problem}[ ] ...  \end{problem}
%% \begin{example}[ ] ...  \end{example}
%% \begin{remark}[ ] ...  \end{remark}
%% \begin{proof} ...  \end{proof}
%%
 \section{Conclusion}
In this work a new association rules mining algorithm based on penguins search optimization 
algorithm (Pe-Arm) has been proposed. The incorporation of a new overlapping measure  into the main method of Pe-Arm to evaluate the amount of overlapping between generated rules in order to generate a set of consistent rules. 
Pe-Arm has been compared with a set of well known meta-heuristics for association rule mining, by different criteria such as the computational time, the statistical measure (confidence and support) and finally with the coverage measure, which proved the efficiency of the proposed Pe-ARM. Different data sets have been used in the comparison, firstly is the standard data is used to evaluate the approach with those datasets that have been used by most of the association rule mining algorithm. Secondly, with a biological datasets, the use of biological data sets allows us to validate the approach with data that contain a huge number of association between item sets which is the specificity of the gene expression data. The gene expression datasets is made for that raison like extracting rules between conditions of genes. Currently, we are investigating with the improvement of the rules representation to develop new way to represent large datasets because we lose a lot of time on measure calculation.

>>>>

\title{Penguins Search Optimisation Algorithm for Association Rules Mining}

@@

\address[a]{Department of Computer Science, University of Hull, Hull HU6 7RX, UK}

\address[b]{ Badji Mokhtar-Annaba University, P.O. Box 12, 23000 Annaba, Algeria}

@@

Association Rules Mining (ARM) is one of the most popular and well-known approaches for the decision-making process. All existing ARM algorithms generate a very large number of association rules with high overlapping. To deal with this issue, we propose a new ARM approach based on penguins search optimisation algorithm (Pe-ARM for short). Moreover, an efficient measure is incorporated into the main process to evaluate the amount of overlapping among the generated rules. Indeed, before the evaluation of the rules, each penguin calculates the amount of overlapping between its rule and all accepted rules. Then, the given rule is rejected without fitness evaluation if at least the amount of overlapping for this rule and one rule from the set of accepted rule is greater than the maximum accepted overlap. The proposed approach also ensures a good diversification over the whole solutions space thanks to the multi-groups behaviour of penguins. Furthermore, the search process is based on the oxygen reserve of each penguin, which permit to accelerate the search of a good solutions in a reasonable time and to save energy in the disagreeable regions. To demonstrate the effectiveness of the suggested approach, several experiments have been carried out on different data-sets and specifically on the biological ones. The results reveal that the proposed approach outperforms the well known ARM algorithms in both execution time and solution quality. 
\end{abstract}

\begin{keyword}
Association Rule Mining; Penguin Search Optimisation Algorithm; Overlap measure; Biological Data-set; ARM.
%% keywords here, in the form: keyword \sep keyword

@@

 Association Rules Mining is one of the most challenging and important tasks in data mining~\cite{1}. 

@@

the problem solving process. In the literature, ARM algorithms are divided into two main categories: exact approaches and 
meta-heuristics based approaches. The exact approaches such as Apriori~\cite{5}, FPGrowth~\cite{6}, DIC~\cite{7}, DHP~\cite{8} and Eclat~\cite{9} aim at extracting all possible association rules. However, these algorithms are high time and space consuming when dealing with very large databases with many items and transactions. 
To reduce the computation time of such algorithms, a second category, metaheuristics-based approaches have been proposed. Examples of such approaches include ARMGA~\cite{10} and G3PARM~\cite{11} for evolutionary algorithms, PSO-ARM~\cite{12}, $ACO_R$~\cite{13} and BSO-ARM~\cite{14} 
for swarm intelligence. These algorithms have the potential to \textbf{give only one part of rules and not all relevant rules}. Therefore, two challenges can be derived, the computation time of exact algorithms by exploring all rules space in one hand and the quality of the extracted rules using meta-heuristics based approaches by taking into account the similar rules in the other hand. \\
The penguins search optimisation algorithm (PeSOA) \cite{15} is a recent nature inspired approach based on the collaborative hunting strategies of penguins. The penguins synchronise their dives to reduce the expenditure of energy in the hunting process. This strategy is used in the algorithm to help converging quickly towards an optimal solution. The diversification strategy of the penguins search algorithm allows penguins to explore  the whole solution space efficiently. The search process of penguins is based on the reserve of oxygen which allows penguins to decide whether to accelerate or decelerate while exploring the search space and also to decide whether to search or not in a given region. This has motivated the use of the PeSOA for the association rules mining problem to facilitate efficient exploration of the solution space. We incorporate a new measure to the main process of the penguins search to compute the amount of overlapping between rules to help penguins search process to generate only the non redundant rule with low amount of overlapping among them. \\
In this paper, a new algorithm called Pe-Arm (penguins search optimisation algorithm for association rules mining) is proposed  to find a set of consistent rules with superior quality and a low amount of overlapping among the rules. These set of rules must cover the maximum number of transaction in the database.\\
The remainder of the paper is organised as follows: In the next section, relevant works on association rule mining with diverse applications are presented. 
Then, the penguins search optimisation algorithm is introduced in Section 3, 
followed by the proposed Pe-ARM approach   in section 4. 
The experimental results with both standard and biological data sets are reported in Section 5. 

@@

Particle swarm optimisation algorithm has been applied for association rule mining \cite{12}, 

@@

Other swarm intelligence algorithms have also been applied for ARM problem. First, an Ant colony optimisation \cite{20} worked in health application 

@@

a gravitational emulation local search algorithm, which is a nature inspired optimisation algorithm. Each rule is modeled as a mass, all the masses attract each others using the law of motion. Each iteration is based on the 

@@

Bees Swarm Optimisation (BSO) and Tabu Search (TS). BSO will explore the search space so that it can cover most of its neighbors. 

@@

\section{Penguins Search Optimisation Algorithm }

Penguin search optimisation algorithm (PeSOA) is a new swarm based metaheuristic algorithm which was proposed in \cite{15}. The dietary behaviour of penguins may be explained by the economic reasoning: it comes to a profitable food search activity when the gain of energy is greater than the expenditure required to obtain this gain. Penguins, behaving along the line of foraging predators, must extract information about the time and cost to get food and the energy content of prey in order to choose the course for making their next dive.
 PeSOA is inspired by the penguin’s hunting behaviour and it generally works as follows. The population of penguins locates initial positions, this population is divided on a set of groups, and each group is assigned to a region on the whole solution space. Each penguin then dives and swims under the water for hunting fish while consuming its oxygen reserve. Different forms of the communication between penguins are occasionally performed and the quantities of eaten fish increase. The process is repeated until the specified amount of fish is obtained or the maximum number of iterations is reached. The authors of \cite{15} have shown the PeSOA outruns genetic algorithms and particle swarm optimisation in obtaining better values for benchmark optimisation functions. 

@@

The optimisation of ARM aims at maximizing the

@@

rules. Optimisation algorithm gives only a set of the pertinent

@@

 After each iteration, the penguins communicate with each other the best rule founded  $GLbest$ to converge for the best group, update the oxygen reserve and the number of neighbours  for each penguin. The computation of the health of each group is made to  define the probabilities of improvement in each group, and finally we redistribute the penguins on the new population according the health of the penguins of each group (probabilities). 
\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=15cm]{2.jpg}
\end{center}
\caption{Penguins Search Optimisation Algorithm For Arm}
\label{algo}
\end{figure*}
\section{EXPERIMENTATION AND RESULTS}
We used several evaluation criteria in the experimentation process in order to evaluate the proposed algorithm.
Firstly, the statistical measure is computed which is represented as the average of the confidence and the support of the generated rules. 
Secondly, the run time of each approach is determined. \\
The last measure is the coverage formula which represents the similarities 
between the generated rules according to the number of common transactions. Indeed, 
the rules are similar when they verify together many transactions and dissimilar where
they do not verify any transaction \cite{27}. The coverage formula is given as follows:\\
Let $T_{r_i}$ be the set of transactions verified by $r_{i}$, and $n$ is the number of generated rules.\\
\begin{displaymath}
Coverage =   \frac{\sum_{i=1}^{n}{\sum_{j=1}^{n} {\eta{(r_{i},r_{j})}}}}{n~(n-1)} \\
\end{displaymath}
\begin{displaymath}
\eta{(r_{i},r_{j})} =
\left\lbrace
\begin{array}{ccc}
0          &  i = j & \\
\left|T_{r_{i}} \cup T_{r_{j}}\right| - \left|T_{r_{i}} \cap T_{r_{j}}\right|  & \mbox{Otherwise}  
\end{array}\right.
\end{displaymath}

@@

%% \begin{theorem}[ ] ...  \end{theorem}
%% \begin{definition}[ ] ...  \end{definition}
%% \begin{axiom}[ ] ...  \end{axiom}
%% \begin{conjecture}[ ] ...  \end{conjecture}
%% \begin{proposition}[ ] ...  \end{proposition}
%% \begin{corollary}[ ] ...  \end{corollary}
%% \begin{lemma}[ ] ...  \end{lemma}
%% \begin{problem}[ ] ...  \end{problem}
%% \begin{example}[ ] ...  \end{example}
%% \begin{remark}[ ] ...  \end{remark}
%% \begin{proof} ...  \end{proof}
%%
 \section{Conclusion}
In this work a new association rules mining algorithm based on penguins search optimisation 
algorithm (Pe-Arm) has been proposed. The incorporation of a new overlapping measure  into the main method of Pe-Arm to evaluate the amount of overlapping between generated rules in order to generate a set of consistent rules. 
Pe-Arm has been compared with a set of well known meta-heuristics for association rule mining, by different criteria such as the computational time, the statistical measure (confidence and support) and finally with the coverage measure, which proved the efficiency of the proposed Pe-ARM. Different data sets have been used in the comparison, firstly is the standard data is used to evaluate the approach with those datasets that have been used by most of the association rule mining algorithm. Secondly, with a biological datasets, the use of biological data sets allows us to validate the approach with data that contain a huge number of association between item sets which is the specificity of the gene expression data. The gene expression datasets is made for that raison like extracting rules between conditions of genes. Currently, we are investigating with the improvement of the rules representation to develop new way to represent large datasets because we lose a lot of time on measure calculation.
